---
title: "Homework 0"
output: pdf_document
---

## Name: Put your name here

## I worked with: 

**Click the "Knit" button in RStudio to knit this file to a pdf.**

--------------------------------------------

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      size = "small", 
                      collapse = TRUE, 
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE,
                      error = TRUE) # document will compile even in case of error

```


```{r}
# load necessary packages 
library(ggplot2)
library(dplyr)
library(gapminder)
library(babynames)
```


**Problem 1: Gapminder**

The `gapminder` package includes a data frame called `gapminder`, containing information about different countries from 1952 to 2007. We are interested in the year 2007 and we are going to save this in a new data frame `gapminder_2007` below. We use data wrangling using code from the `dplyr` package. We will learn about in detail later this later.


```{r}
gapminder_2007 <- gapminder %>% 
  filter(year == 2007)
```


Run `View(gapminder_2007)` in your console to explore this data. An alternative method for exploring a data frame is by using the `glipmse()` function:

```{r}
glimpse(gapminder_2007)
```

The function below plots the life expectancy Vs GDP per capita (in USD) of all the countries in the `gapminder` dataset. You can see these individual points color-coded by continent and the size of the points correspond to the population of the country they correspond to. We will dissect this function in detail in the coming weeks.


```{r, fig.cap= "A nice visualization", tidy=FALSE}
ggplot(data = gapminder_2007, 
       mapping = aes(x = gdpPercap, y = lifeExp, size = pop, color = continent)) +
  geom_point() +
  # Note: this was not required for this problem set:
  labs(x = "GDP per capita (in USD)", 
       y = "Life expectancy", 
       size = "Population", 
       color = "Continent", 
       title = "Gapminder data",
       subtitle = "2007 values")
```


a. Comment on the two biggest green circles that you see in the Figure 1 above.

*Answer:*

\vspace*{0.2in}


**Problem 2: Babynames**

The `babynames` package provides data about the popularity of individual baby names from the US Social Security Administration. Data includes all names used at least 5 times in a year beginning in 1880.

In the code chunk below, we extract the data corresponding to `Aimee` or `Sammy` from `babynames` dataset and store it in a new dataset `babynames_Aimee_Sammy`.  

```{r}
babynames_Aimee_Sammy <- babynames %>%
  filter(name == "Aimee" | name == "Sammy")
```

a. What do you see in Figure 2 below? We will learn the functions in detail later.

```{r, fig.cap="A nice comparison", tidy=FALSE}
ggplot(babynames_Aimee_Sammy, aes(x=year, y=prop, col=sex)) +
  geom_line() +
  facet_wrap(~name) +
  labs(x = "Year", y = "Proportion", color = "Sex", title = "Comparison of Aimee and Sammy")
```

*Answer:*


